---
title: "Prediction Assignment"
author: "Juliaty Tio"
date: "November 5, 2016"
output: html_document
---

##Background
Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes. Participants were supervised by an experienced weight lifter to make sure the execution complied to the manner they were supposed to simulate. The exercises were performed by six male participants aged between 20-28 years, with little weight lifting experience. We made sure that all participants could easily simulate the mistakes in a safe and controlled manner by using a relatively light dumbbell (1.25kg). The goal of your project is to predict the manner in which they did the exercise. 
Read more: http://groupware.les.inf.puc-rio.br/har#ixzz4P4AW29ar

##Data Preparation

Firstly, we load the training and testing data from the physical location into r enviroment. 
```{r,message=FALSE }
library(caret)
library(randomForest)

trainingset <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""));
testingset <-read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!", ""));
```

### Clean up data
Before we use the model, we need to clean the data first. Firstly we omit the columns that contain na/null data, then we remove the timestamp columns and all other character type columns. This procedure was done to both training and testing set.
```{r}
#training set cleanup
training <-trainingset[,colSums(is.na(trainingset)) == 0]
col.name <- names(training)
col.name <- col.name[-grep("*timestamp",col.name)]    
training<-training[,col.name];
training<-training[,-c(1:3)];

#testing set cleanup
testing <-testingset[,colSums(is.na(testingset)) == 0]
col.name <- names(testing)
col.name <- col.name[-grep("*timestamp",col.name)]    
testing<-testing[,col.name];
testing<-testing[,-c(1:3)];
names(testing)
```


### Partitioning the data
We split the training data into 75% for training and 25% for testing, as the provided testing data will be use for validation. 
```{r}
inTrain <- createDataPartition(training$classe,p=0.75,list=FALSE);
train.data <- training[inTrain,];
test.data <- training[-inTrain,];
```

##Model Fitting
We will use random forest method with cross validation with 15 fold as the train control to increase the accuracy. 
```{r}
set.seed(11115);
ctrl<-trainControl(method="cv",number=15);
fit.ctrl <- train(data=train.data,train.data$classe~.,method="rf",trControl=ctrl);
fit.ctrl
```
From the result we can see the accuracy is in 99.7%, next we will see the accuracy from the test set that we have defined before. 
 
##Prediction and result
```{r}
pred.fit.ctrl <- predict(fit.ctrl,test.data);
confusionMatrix(test.data$classe,pred.fit.ctrl)

```

From the result, we can see the out of sample error which is only around 0.3% chance of error, so the accuracy is quite good. Below script will generate the files for the testing prediction result. 

```{r}

result <- predict(fit.ctrl,testing)
write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

write_files(result)
```
 





